### 2015-09-15 Irregular-occ/Irregular-occ.tex
.

\abstract
{The irregular satellites of the giant planets are believed, due to their orbital configurations, to not have been formed where they are located, but they were probably captured in the early solar system. To constrain their origin location, it's important to know their physical parameters, such as size, density and albedo. These informations would to understand better their densities and consequently their composition. The best ground-based technique to obtain these parameters is the observation of stellar occultations by these objects.}
{We aimed to predict stellar occultation for the seven biggest irregular satellites of Jupiter: Himalia, Elara, Pasiphae, Carme, Lysithea, Sinope and Ananke; for the irregular satellite of Saturn: Phoebe; and the satellites of Neptune: Nereid and Triton.}
{We identified stars candidates to stellar occultation by the irregular satellites in the UCAC4 catalogue and a catalogue of stars in the sky-path of Neptune obtained from observations made with the ESO2p2/WFI (2.2 m Max-Planck ESO telescope with the Wide Field Imager) instrument. These catalogue were crossed with the JPL ephemeris of the satellites using the methodology applied to identify stellar occultation by TNOs by \cite{Assafin2012}. The ephemeris were corrected of offsets using an numerical integration \textcolor{red}{(Here we may write something about after the work of Laurène)} of the positions obtained by \cite{GomesJunior2015} to improve the predictions.}
{We managed to obtain \textcolor{red}{588} candidates of stellar occultations between the period of January, 2015 and December, 2017. Tests were develop to obtain offsets for the star and satellite for the events of Himalia in March 03, 2015 and Elara in March 30, 2015. The observations were made in the same field to minimize errors and the offsets were used to correct the predictions and evaluate them.}
{The comparisons between the nominal predictions, the corrected predictions using the numerical integration and the observation tests show a good agreement. This reveals that observe an stellar occultation by these objects, at least for Himalia, is quite possible.}

@@

In this work, we propose to study these objects as possible representatives of the small TNOs' population. The basis for this hypothesis is discussed in Section \ref{Sec: Rationale}. TNOs are highly interesting objects that due to their large heliocentric distances (low temperatures) may be highly preserved having their physical properties similar to those they had when they were formed, then providing history and evolution of the outer solar system \citep{Camargo2013}. This is even more true for the smaller objects, since in principle larger sizes favour physical differentiation processes in the body and vice-versa. However, due to the distance, the smaller TNOs from this region are more difficult to observe. A clever way to overcome this difficulty is to study  the much closer irregular satellites, under the hypothesis that they share a common origin with the small TNOs' population.

In order to obtain precise fundamental physical parameters like size, shape and albedo for the irregular satellites (and thus, in a sense, to small-sized TNOs), we will make use of stellar occultations, which provide  more accurate results than other ground-based techniques \citep{Sicardy2011, Ortiz2012, Braga-Ribas2014}.

Up to date no stellar occultation by irregular satellite is found in the literature. Since their estimated sizes are very small (see table \ref{Tab: satellite-diameter}), prediction of the exact location and instant where the shadow will cross the Earth demands a good accurate ephemeris and star positions. For instance, Himalia, supposedly the largest irregular satellite of Jupiter has an estimated size of 150 km \citep{Porco2003}, which is equivalent to an apparent size of about 40 mas (milliarcseconds).Thus, even in the best cases, the overall budget of ephemeris and star position errors must be around 40 mas in order that we get good chances of observing a stellar occultation.

@@

Unlikely stellar occultations by TNOs, which has been proved effective, the observation of an occultation by a irregular satellite is more favorable. They have been known for more time and observed in many orbital period having better ephemeris than TNOs. Moreover, the irregular satellites are closer to Earth which means a minor error in kilometers.

\cite{GomesJunior2015} obtained 6523 suitable positions for 18 irregular satellites between 1992 and 2014 with an estimated error in the positions of about 60 to 80 mas. For some satellites the number of positions obtained is comparable to the number used in the numerical integration of orbits by the JPL \citep{Jacobson2012}. They pointed out that the ephemeris of the irregular satellites have systematic errors that may reach 200 mas for some satellites. For an object at the distance of Jupiter, this represents an error bigger than 700 km in the shadow path. Using the offsets obtained by \cite{GomesJunior2015} it is possible to correct the ephemeris and better predict stellar occultations for these objects.

We present in this paper stellar occultation predictions for the 7 major irregular satellites of Jupiter (Himalia, Elara, Pasiphae, Lysithea, Carme, Ananke and Sinope), Phoebe from Saturn and Nereid and Triton from Neptune. In the section \ref{Sec: Rationale} we explore the scientific rationale for the study of the irregular satellites and the possibility of having a common origin with TNOs. In section \ref{Sec: Offset} we show the correction made to the ephemeris of these satellites for better predicting the stellar occultations. In section \ref{Sec: predictions}, we present the predictions of the stellar occultations by irregular satellites and how they were made. Some tests made to confirm the predictions are presented in section \ref{Sec: testes} and the conclusions are given in section \ref{Sec: conclusions}.

@@

As explicited in section \ref{Sec: introducao}, there is no consensus on a single model about the region where the irregular satellites were formed. \cite{Cuk2004} showed that the progenitor of the Himalia group may have origin in heliocentric orbits similar to the Hilda asteroid group. \cite{Sheppard2005} stated that the irregular satellites may be some of the objects that were formed within the giant planets region. \cite{Sheppard2005} and \cite{Jewitt2007} also expose the possibility for the irregular satellites having their origin as comets or similar to the TNOs.

Phoebe, the most studied irregular satellite, was identified by \cite{Clark2005} that its surface is probably covered by material of cometary origin. It was also stated by \cite{Johnson2005} that if the porosity of Phoebe would be of about 15\%, which the authors believe it is possible, Phoebe would have an uncompressed density similar to those of Pluto and Triton.

Triton is a special satellite. Its orbit is retrograde and inclined, but quasi-circular and very close to the planet compared to the irregular ones. Because Triton's orbit is very small and its precession is not dominated by Solar perturbations, Triton is frequently excluded of the irregular satellites' class, but studied together by many authors \citep{Sheppard2005, Jewitt2007}.

Similarly to the irregular satellites, Triton was probably capture in the early Solar System and may have a similar origin as the TNOs \citep{Agnor2006}. Differently, Triton is bigger than the irregular satellites by an order of magnitude and has an atmosphere. The main objective to study Triton by stellar occultation is to understand the evolution of its atmosphere due to Triton's complicated and extreme seasonal cycle \citep{McKinnon2007}.

@@

\begin{figure}
\begin{centering}
\includegraphics[scale=0.2]{figures/Himalia_2015-10-23T07:54:47.png} 
\caption{Occultation map for Himalia regarding to the second event sampled in Table \ref{Tab: occ-list}}
\label{Fig: ocultacao}
\end{centering}
\end{figure}

>>>>

\abstract
{Due to their orbital configurations, it is common belief that the irregular satellites were probably captured by the giant planets in the early solar system. It is important to know their physical parameters, such as size and albedo, to trace back their true origin. The best ground-based technique to obtain such parameters is the observation of stellar occultations by these objects.}
{We aimed to predict stellar occultations for the seven biggest irregular satellites of Jupiter: Himalia, Elara, Pasiphae, Carme, Lysithea, Sinope and Ananke; for the irregular satellite of Saturn: Phoebe; and the satellites of Neptune: Nereid and Triton.}
{We identified candidates to stellar occultations by the irregular satellites from the UCAC4 catalogue and from a catalogue of stars in the sky-path of Neptune obtained from observations made with the ESO2p2/WFI (2.2 m Max-Planck ESO telescope with the Wide Field Imager) instrument. These catalogues were crossed with the ephemeris of the satellites to identify stellar occultations. The ephemeris were generated by a numerical integration \textcolor{red}{(Here we may write something about after the work of Laurène)} that used the positions obtained by \cite{GomesJunior2015} to improve the predictions.}
{We managed to obtain \textcolor{red}{588} candidates of stellar occultations between the period of January, 2015 and December, 2017. We made observational tests for the predictions of the events of Himalia in March 03, 2015 and Elara in March 30, 2015. The observations were made in the same field to minimize errors and the obtained relative satellite-star positions were used to evaluate the predictions.}
{The comparisons between the predictions and the observation tests show a good agreement. We discuss how the successful observation of a stellar occultation by these objects is quite possible.}

@@

In order to obtain precise fundamental physical parameters like size, shape and albedo for the irregular satellites and thus to contribute to the study of their origin, we will make use of stellar occultations, which provide  more accurate results than other ground-based techniques \citep{Sicardy2011, Ortiz2012, Braga-Ribas2014}.

No observation of a stellar occultation by an irregular satellite was published up to date. Since their estimated sizes are very small (see Table \ref{Tab: satellite-diameter}), this may have prevented earlier tries. But, in fact, given their distances to us, current ephemeris and star positions are already good enough for the prediction of the exact location and instant where the shadow of the occultation will cross the Earth. For instance, Himalia, supposedly the largest irregular satellite of Jupiter has an estimated size of 150 km \citep{Porco2003}, which is equivalent to an apparent size of about 40 mas (milliarcseconds).Thus, in this case, the overall budget of ephemeris and star position errors must be around 40 mas in order that we get good chances of observing a stellar occultation, which is quite feasible today.

@@

Unlike stellar occultations by TNOs, which nevertheless have been proved effective, the observation of an occultation by a irregular satellite is, in principle, more favorable. The orbits of their host planets are well known and these satellites have been observed completing already many orbital periods around them, thus presenting better ephemeris than TNOs. Moreover, the irregular satellites are closer to Earth which means a minor location error in kilometers.

\cite{GomesJunior2015} obtained 6523 suitable positions for 18 irregular satellites between 1992 and 2014 with an estimated error in the positions of about 60 to 80 mas. For some satellites the number of positions obtained is comparable to the number used in the numerical integration of orbits by the JPL \citep{Jacobson2012}. They pointed out that the ephemeris of the irregular satellites have systematic errors that may reach 200 mas for some satellites. For an object at the distance of Jupiter, this represents an error bigger than 700 km in the shadow path. Using the positions obtained by \cite{GomesJunior2015} it is possible to improve the ephemeris and better predict stellar occultations for these objects.

We present in this paper stellar occultation predictions for the 7 major irregular satellites of Jupiter (Himalia, Elara, Pasiphae, Lysithea, Carme, Ananke and Sinope), Phoebe from Saturn and Nereid and Triton from Neptune. In the section \ref{Sec: Rationale} we explore the scientific rationale for the study of the irregular satellites and the possibility of having a common origin with TNOs. In section \ref{Sec: Offset} we show the \textcolor{red}{numerical integration to improve the ephemeris}. In section \ref{Sec: predictions}, we present the predictions of the stellar occultations by irregular satellites and how they were made. Some tests made to confirm the predictions are presented in section \ref{Sec: testes} and the conclusions are given in section \ref{Sec: conclusions}.

@@

There is no consensus for a single model explaining where the irregular satellites were formed. \cite{Cuk2004} showed that the progenitor of the Himalia group may have originated in heliocentric orbits similar to the Hilda asteroid group. \cite{Sheppard2005} stated that the irregular satellites may be some of the objects that were formed within the giant planets region.

Phoebe is the most studied irregular satellite. \cite{Clark2005} suggest that its surface is probably covered by material of cometary origin. It was also stated by \cite{Johnson2005} that if the porosity of Phoebe would be of about 15\%, which the authors believe it is possible, Phoebe would have an uncompressed density similar to those of Pluto and Triton.

\cite{Sheppard2005} and \cite{Jewitt2007} also expose the possibility for the irregular satellites having their origin as comets or similar to the TNOs. TNOs are highly interesting objects that due to their large heliocentric distances (low temperatures) may be highly preserved having their physical properties similar to those they had when they were formed, then providing history and evolution of the outer solar system \citep{Camargo2014}. This is even more true for the smaller objects, since in principle larger sizes favour physical differentiation processes in the body and vice-versa. However, due to the distance, the smaller TNOs from this region are more difficult to observe. A clever way to overcome this difficulty is to study  the much closer irregular satellites, under the hypothesis that they share a common origin with the small TNOs' population.

Triton is a special satellite. Its orbit is retrograde and inclined, but quasi-circular and very close to the planet compared to the irregular ones. Because Triton's orbit size is very small and its precession is not dominated by Solar perturbations, Triton is frequently excluded from the irregular satellites' class, but studied together by many authors \citep{Sheppard2005, Jewitt2007}.

Similarly to the irregular satellites, Triton was probably captured in the early solar system and may have a similar origin as the TNOs \citep{Agnor2006}. Differently, Triton is bigger than the irregular satellites by an order of magnitude and has an atmosphere. The main motivation here to study Triton by stellar occultations is to understand the evolution of its atmosphere due to Triton's complicated and extreme seasonal cycle \citep{McKinnon2007}.

@@

\begin{figure*}
\begin{centering}
\includegraphics[scale=0.4]{figures/Himalia_2015-10-23T07:54:47.png} 
\caption{Occultation map for Himalia regarding to the second event sampled in Table \ref{Tab: occ-list}}
\label{Fig: ocultacao}
\end{centering}
\end{figure*}
